---
title: "Tutoriel : utiliser une instance Postgres avec R"
author: "Pierre Lamarche"
date: "01/06/2020"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectif du tutoriel

Ce tutoriel illustre la connexion depuis `R` à une instance Postgres qui contient les données du Recensement de Population 2016, disponibles sur le [site de l'Insee](https://www.insee.fr/fr/statistiques/4171558). L'instance en question a été déployée sur la plateforme Onyxia du SSPCloud, et n'est accessible que depuis cette plateforme.

On montera quelques usages que l'on peut faire à partir de cette instance et en quoi combiner `R` et Postgres est utile.

## Pourquoi utiliser une base de données

Les données du Recensement de Population sont assez volumineuses ; pour les données de 2016, elles représentent environ 63 Go. Il est par conséquent illusoire de vouloir travailler avec l'ensemble de ces données chargées simultanément en mémoire avec `R`. La seule table `logement` contient 25 millions de lignes et peut déjà représenter une difficulté pour certaines machines. Ces contraintes d'espace tombent avec un système de base de données. 

## Comment utiliser une base de données avec `R`

### Se connecter à la base Postgres


To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



